---
title: "Analysis of Credit Card Default"
author: "Hafid Pradipta"
date: "April 17, 2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}

library(readxl)
library(tidyverse)
library(ggplot2)
library(leaps)
library(FactoMineR)
library(factoextra)
library(MASS)
library(biotools)
library(randomForest)
library(gridExtra)
library(MVN)
library(rcompanion)
library(ROCR)
credit <- read_excel("C:/Users/Hafid Pradipta/OneDrive/ITEC 621/credit card default/creditdefault.xls")
credit
```

```{r tidy up the data,message=FALSE}
#tidy up the data
credit2 <- credit
credit <- credit %>%
  mutate(SEX=ifelse(credit$SEX==1,"male","female"))
credit$MARRIAGE <- as.factor(credit$MARRIAGE)
credit$MARRIAGE <- credit$MARRIAGE %>% 
  recode(`1` = "Married",
         `2` = "Single",
         `3` = "Others")

credit$EDUCATION <- as.factor(credit$EDUCATION)

credit$EDUCATION <-  credit$EDUCATION %>% 
  recode(`1`= "Gradschool",
         `2` = "University",
         `3` = "Highschool",
         `4` = "Others",
         `5` = "Others",
         `6` = "Others")
colnames(credit)[25] <- ("defaultnm")
credit <- credit %>% 
  mutate(defaultnm = ifelse(credit$defaultnm==1, "default","nondefault"))

hist(credit$BILL_AMT1)
```


```{r convert NT to USD}
#conver NTdollar to US dollar
credit <- credit %>% 
  mutate(limbalusd =LIMIT_BAL*0.034)

credit <- credit %>% 
  mutate(flimbalusd = cut(limbalusd, c(0, 5000, 15000, 20000 ,25000,30000, Inf), right = FALSE,
                          labels = c("0-5k","5-15k","15-20k","20-25k","25-30k","30k")))

credit <- credit %>%
  mutate(billamt1 = BILL_AMT1*0.034,
         billamt2 = BILL_AMT2*0.034,
         billamt3 = BILL_AMT3*0.034,
         billamt4 = BILL_AMT4*0.034,
         billamt5 = BILL_AMT5*0.034,
         billamt6 = BILL_AMT6*0.034,
         payamt1 = PAY_AMT1*0.034,
         payamt2 = PAY_AMT2*0.034,
         payamt3 = PAY_AMT3*0.034,
         payamt4 = PAY_AMT4*0.034,
         payamt5 = PAY_AMT5*0.034,
         payamt6 = PAY_AMT6*0.034)

# Normality Test
n <- length(credit$defaultnm)
cnum <- credit[,c(26, 28:39)]
ZZ <- sample(n,5000)
cnumnorm <- mvn(data = cnum[ZZ,], univariateTest = "SW", univariatePlot = TRUE)
cnumnorm$univariateNormality
scnum <- scale(cnum, scale = TRUE)
scnumnorm <- mvn(data = scnum[ZZ,], univariateTest = "SW", univariatePlot = TRUE)
ccnum <- scale(cnum, scale = FALSE, center = TRUE)
ccnumnorm <- mvn(data = scnum[ZZ,], univariateTest = "SW")
nresult <- tibble(names = colnames(cnum),Skew=cnumnorm$Descriptives$Skew, Kurt=cnumnorm$Descriptives$Kurtosis,
                  ScaleSkew=scnumnorm$Descriptives$Skew, ScaleKurt=scnumnorm$Descriptives$Kurtosis,
                  CenterSkew=ccnumnorm$Descriptives$Skew, CenterKurt=ccnumnorm$Descriptives$Kurtosis)


#Tukey Transformation
a <- transformTukey(cnum$limbalusd[ZZ],plotit=FALSE)#0.25
ab1 <- transformTukey(cnum$billamt1[ZZ],plotit=FALSE)#1
ab2 <- transformTukey(cnum$billamt2[ZZ],plotit=FALSE)#1
ab3<- transformTukey(cnum$billamt3[ZZ],plotit=FALSE)#1
ab4 <- transformTukey(cnum$billamt4[ZZ],plotit=FALSE)#1
ab5<- transformTukey(cnum$billamt5[ZZ],plotit=FALSE)#1
ab6 <- transformTukey(cnum$billamt6[ZZ],plotit=FALSE)#1
ap1 <- transformTukey(cnum$payamt1[ZZ],plotit=FALSE)#0.3
ap2 <- transformTukey(cnum$payamt2[ZZ],plotit=FALSE)#0.25
ap3 <- transformTukey(cnum$payamt3[ZZ],plotit=FALSE)#0.25
ap4 <- transformTukey(cnum$payamt4[ZZ],plotit=FALSE)#0.25
ap5 <- transformTukey(cnum$payamt5[ZZ],plotit=FALSE)#0.25
ap6 <- transformTukey(cnum$payamt6[ZZ],plotit=FALSE)#0.25

#transformation to make it more normal
cnumt <- data.frame(cnum[,1]^0.25,
          cnum[,2:7],
          cnum[,8:13]^0.25)
cnumtnorm <- mvn(data = cnumt[ZZ,],univariateTest = "SW")
cnumtnorm$univariateNormality
creditt <- credit
creditt[,26:39] <- cnumt
creditt$flimbalusd <- credit$flimbalusd
names(creditt)
```

```{r conversion to long dataset,message=FALSE }

abc <- credit %>% gather(paste0("billamt",1:6), key = "billperiod", value = "billamount")
abc <- abc %>% gather(paste0("payamt",1:6), key = "payperiod", value = "payamount")
longdata <- abc[,c(25,28:31)]

credit[,7:12] <- credit[,7:12]+2
def <- credit %>% gather(PAY_0,paste0("PAY_",2:6), key = "payperiod", value = "paypunc")
longdata2 <- def[,34:35]

```



```{r default visualization}
#data visualization for default group. 
default <- credit %>% filter(defaultnm == "default")

d.sex<- default %>% group_by(SEX) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x =SEX, y = count), label = count)+
  geom_bar(stat="identity", fill = "midnightblue")+
  geom_text(aes(label = round(count/sum(count),2)), size=3, vjust = -0.25)+
  xlab("Gender")+
  ylab("Number of People")+
  theme(axis.text.x = element_text(angle = 325, hjust = 0))+
  ggtitle("Default on Gender")

d.edu <- default %>% group_by(EDUCATION) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x =EDUCATION, y = count), label = count)+
  geom_bar(stat="identity",fill = "midnightblue")+
  geom_text(aes(label = round(count/sum(count),2)), size = 3, vjust = -0.25)+
  ylab("Number of People")+
  theme(axis.text.x = element_text(angle = 325, hjust = 0))+
  ggtitle("Default on Education")

d.mar <-default %>% group_by(MARRIAGE) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x =MARRIAGE, y = count))+
  geom_bar(stat="identity", fill = "midnightblue", position = position_stack())+
  geom_text(aes(label = round(count/sum(count),2)), size = 3, vjust = -0.25)+
  ylab("Number of People")+
  xlab("Marital Status")+
   theme(axis.text.x = element_text(angle = 325, hjust = 0))+
  ggtitle("Default on Marital")

d.sex
d.edu
d.mar
```

```{r EDA,message=FALSE}
#cxredit limit visualization
credit %>% 
  group_by(flimbalusd) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x =flimbalusd, y = count), label = count)+
  geom_bar(stat="identity", fill = "midnightblue")+
  geom_text(aes(label = round(count/sum(count),2)), size = 5, vjust = -0.25)+
  theme(legend.position="none")+
  xlab("Limit Category")+
  ylab("Amount of limit in USD")+
  ggtitle("Credit Limit in USD")

```

We split the limit balance per 5k and based on the graph, most of the customers has limit balance between $5000 - %15000 which account for 95% of all customers. 

```{r average payment and average bill,message=FALSE}

#average bill amount, bill paid and lateness of the payment visualization. 

badef <- longdata %>% 
  filter(defaultnm=="default") %>% 
  group_by(billperiod) %>% 
  summarise(average= mean(billamount))
bandef <- longdata %>% 
  filter(defaultnm=="nondefault") %>% 
  group_by(billperiod) %>% 
  summarise(average= mean(billamount))

diffba <- badef %>% left_join(bandef, by = "billperiod")
colnames(diffba)[2:3] <- c("default","nondefault")
diffba$billperiod <- diffba$billperiod %>% 
  recode("billamt1"="bperiod1",
         "billamt2"="bperiod2",
         "billamt3"="bperiod3",
         "billamt4"="bperiod4",
         "billamt5"="bperiod5",
         "billamt6"="bperiod6")


diffba %>% gather(default, nondefault, key= "Type", value = "amount") %>% 
  ggplot(aes(x= billperiod, y=amount))+
  geom_point(aes(color = Type, size = amount))+
  geom_text(aes(label = round(amount,2)), vjust = -0.5)+
  guides(size = FALSE)+
  xlab("Period of Payment")+
  ylab("Average of Bill in USD")+
  ggtitle("Average Bill per Period")


ppdef <- longdata %>% 
  filter(defaultnm=="default") %>% 
  group_by(payperiod) %>% 
  summarise(average = mean(payamount))

ppnondef <- longdata %>% 
  filter(defaultnm=="nondefault") %>% 
  group_by(payperiod) %>% 
  summarise(average = mean(payamount))


diffpp <- ppdef %>% left_join(ppnondef, by = "payperiod")
colnames(diffpp)[2:3] <- c("default","nondefault")
########################################################################
diffpp$payperiod <- as.factor(diffpp$payperiod)
diffpp$payperiod <- diffpp$payperiod %>% 
  recode("payamt1"="pperiod1",
         "payamt2"="pperiod2",
         "payamt3"="pperiod3",
         "payamt4"="pperiod4",
         "payamt5"="pperiod5",
         "payamt6"="pperiod6")

diffpp %>% gather(default, nondefault, key= "Type", value = "amount") %>% 
  ggplot(aes(x= payperiod, y=amount))+
  geom_point(aes(color = Type, size = amount))+
  guides(size = FALSE)+
  geom_text(aes(label = round(amount,2)), vjust = -0.5)+
  xlab("Period of Payment")+
  ylab("Average of Payment in USD")+
  ggtitle("Average Payment per Period")


longdata2 %>% 
  group_by(payperiod) %>% 
  summarise(average = mean(paypunc)) %>% 
   ggplot(aes(x= payperiod, y=average))+
  geom_point(aes( size = average))+
  guides(size = FALSE)+
  geom_text(aes(label = round(average,2)), vjust = -0.5)+
  xlab("Period of Payment")+
  ylab("Average of Payment in USD")+
  ggtitle("Average Payment per Period")

```


```{r correlation matrix,message=FALSE}

library(reshape2)
creditnum <- credit[,c(7:12,26,28:39)]
cormat <- round(cor(creditnum),2)
melted_cormat <- melt(cormat)

# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}
# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}
upper_tri <- get_upper_tri(cormat)
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Heatmap
library(ggplot2)

ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))+
  coord_fixed()

####reorder
reorder_cormat <- function(cormat){
  # Use correlation between variables as distance
  dd <- as.dist((1-cormat)/2)
  hc <- hclust(dd)
  cormat <-cormat[hc$order, hc$order]
}

cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))+
  coord_fixed()

ggheatmap + 
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 2) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 0),
    legend.position = c(0.6, 0.7),
    legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                               title.position = "top", title.hjust = 0.5))
```


```{r principal component,message=FALSE}
#method1
creditusd <- credit[,c(1:12,25, 26, 28:39)]
credit
creditusd <- creditusd %>% mutate(defaultnm = ifelse (defaultnm == "default", 1, 0))
creditusd$defaultnm <- as.numeric(creditusd$defaultnm)
prc <- prcomp(creditusd[,-c(1,3:5)], scale = TRUE)

seig1 <- prc$sdev[1]^2/sum(prc$sdev^2)
seig2 <- prc$sdev[2]^2/sum(prc$sdev^2)

grid <- tibble(y = as.factor(creditusd$defaultnm), pc1 =prc$x[,1], pc2 =  prc$x[,2])
gs <- sample(nrow(grid), 5000)
grid <- grid %>% 
  mutate(y = ifelse(y == 1, "default", "nondefault"))

grid[gs,] %>% ggplot(aes(x = pc1, y = pc2, color = y))+
  geom_point(alpha = 0.5, size = 2)+
 xlab (paste("Component1", round(seig1,2), "%"))+
ylab (paste("Component2", round(seig2,2), "%"))+
 labs(color = "Status")

```

```{r machine learning,message=FALSE}
#logistic Regression Initial model

mod1 <- glm(defaultnm~., data = creditusd, family = binomial(link="logit"))
summary(mod1)
#defaultnm~ LIMIT_BAL + SEX + Marriage + AGE + PAY_0 + PAY_2 + PAY_3 + billamt1 + payamt1 + payamt2 + payamt4

n <- length(creditusd$defaultnm)
Z <- sample(n,n/10)
c.test <- creditusd[Z,]
mod1.1 <- glm(defaultnm~ LIMIT_BAL + SEX + MARRIAGE + AGE + PAY_0 + PAY_2 + PAY_3 + billamt1 + payamt1 + payamt2 + payamt4 , data = creditusd[-Z,], family = binomial(link = "logit"))
summary(mod1.1)

prob = predict(mod1.1, data.frame(c.test))
yespred <- 1*(prob>0.5)

table(creditusd$defaultnm[Z], yespred)
log.in <- mean( creditusd$defaultnm[Z] == yespred)
log.in

attach(creditusd)
creditusd.prob.test <- predict(mod1.1, creditusd, type ="response")[-Z]
pred <- prediction(creditusd.prob.test,creditusd$defaultnm[-Z] )
perf <- performance(pred, "tpr","fpr")
plot(perf, colorize = TRUE)
auc=performance(pred, "auc") 
c(auc@y.name[[1]], auc@y.values[[1]])
#any kind of threshold will gave the same result
```

```{r variable selection,message=FALSE}


null = glm( defaultnm ~ 1, data=creditusd, family = binomial(link="logit") )
full = glm( defaultnm ~ ., data=creditusd, family = binomial(link="logit") )

#step( null, scope=list(lower=null, upper=full), direction="forward" )

#glm(formula = defaultnm ~ PAY_0 + LIMIT_BAL + PAY_3 + payamt1 + 
 #   billamt1 + MARRIAGE + EDUCATION + payamt2 + billamt3 + PAY_2 + 
  #  SEX + PAY_5 + AGE + payamt4 + payamt5 + payamt6 + payamt3 + 
   # billamt2, family = binomial(link = "logit"), data = creditusd)

#step( full, scope=list(lower=null, upper=full), direction="backward" )
#glm(formula = defaultnm ~ LIMIT_BAL + SEX + EDUCATION + MARRIAGE + 
    #AGE + PAY_0 + PAY_2 + PAY_3 + PAY_5 + billamt1 + billamt2 + 
   # billamt5 + payamt1 + payamt2 + payamt3 + payamt4 + payamt5 + 
   # payamt6, family = binomial(link = "logit"), data = creditusd)


modvs <- glm(formula = defaultnm ~ LIMIT_BAL + SEX + EDUCATION + MARRIAGE +     AGE + PAY_0 + PAY_2 + PAY_3 + PAY_5 + billamt1 + billamt2 +    billamt5 + payamt1 + payamt2 + payamt3 + payamt4 + payamt5 +    payamt6, family = binomial(link = "logit"), data = creditusd)
summary(modvs)

#Education and payamount6 is not significant. 
#PAY_0 + payamt2 + payamt1 + PAY_3 + payamt6 + payamt5 + billamt3 + PAY_2 + billamt1

modvs2 <- glm(formula = defaultnm ~ LIMIT_BAL + SEX  + MARRIAGE +     AGE + PAY_0 + PAY_2 + PAY_3 + PAY_5 + billamt1 + billamt2 +    billamt5 + payamt1 + payamt2 + payamt3 + payamt4 + payamt5 , family = binomial(link = "logit"), data = creditusd[-Z,])

prob = predict(modvs2, data.frame(c.test))
yespred <- 1*(prob>0.5)
table(creditusd$defaultnm[Z], yespred)
log.vs <- mean( creditusd$defaultnm[Z] == yespred)
log.vs
#logistic regression predict 78% correctly



#TPR = rep(0,100)
#FPR = rep(0,100)
#for(k in 1:100){
#  fit = glm(formula = defaultnm ~ limbalusd + SEX  + MARRIAGE +     AGE + PAY_0 + PAY_2 + PAY_3 + PAY_5 + billamt1 + billamt2 +    billamt5 + payamt1 + payamt2 + payamt3 + payamt4 + payamt5 , family = binomial(link = "logit"), data = creditusd[-Z,])
#  prob  = predict(fit, data.frame(creditusd[Z,]), type = "response")
 # Yhat = 1*(prob > k/100)
  #TPR[k] = sum(Yhat ==1 & defaultnm==1)/ sum(defaultnm==1)
  #FPR[k] = sum(Yhat ==1 & defaultnm==0)/ sum(defaultnm==0)
#}
#plot(FPR, TPR, xlab="False positive rate", ylab="True positive rate", main="ROC curve")



creditusd.prob.test <- predict(modvs2, creditusd, type ="response")[-Z]
pred <- prediction(creditusd.prob.test,creditusd$defaultnm[-Z] )
perf <- performance(pred, "tpr","fpr")
plot(perf, colorize = TRUE)
auc=performance(pred, "auc") 
c(auc@y.name[[1]], auc@y.values[[1]])

```



##Linear Discriminant Analysis

```{r lda,message=FALSE}

lda.fit <- lda(defaultnm~., data= creditusd, CV = TRUE)
table(creditusd$defaultnm, lda.fit$class)
lda.in <- mean(creditusd$defaultnm == lda.fit$class)
lda.in
```



```{r lda vs,message=FALSE}
lda.fitvs <- lda( defaultnm ~ LIMIT_BAL + SEX  + MARRIAGE + AGE + PAY_0 + PAY_2 + PAY_3 + PAY_5 + billamt1 + billamt2 +    billamt5 + payamt1 + payamt2 + payamt3 + payamt4 + payamt5 , data = creditusd, CV = TRUE)
table(creditusd$defaultnm, lda.fitvs$class)
lda.vs <- mean(creditusd$defaultnm == lda.fitvs$class)
lda.vs
```



```{r qda,message=FALSE}

###qda for all variable can't be done due to rank deficiency in one group. It means that the data has too many multicollinearity. 

qda.fitvs <- qda( defaultnm ~ LIMIT_BAL + SEX  + MARRIAGE + AGE + PAY_0 + PAY_2 + PAY_3 + PAY_5 + billamt1 + billamt2 +    billamt5 + payamt1 + payamt2 + payamt3 + payamt4 + payamt5 , data = creditusd, CV = TRUE)
table(defaultnm, qda.fitvs$class)
qda.vs<- mean(defaultnm == qda.fitvs$class,na.rm=TRUE)
qda.vs
```



```{r KNN,message=FALSE}
creditusd2 <- creditusd
creditusd2[,3:5] <- credit2[,3:5]
creditusd2 <- creditusd2[,-1]
creditusd2$defaultnm <- as.factor(creditusd2$defaultnm)
n = length(creditusd2$defaultnm)
Z = sample( n, n/10)
X = model.matrix(defaultnm~., creditusd2)
Y = creditusd2$defaultnm
c.train = creditusd2[-Z,]
c.test = creditusd2[Z,]
x.train <- X[-Z,]
x.test <- X[Z,]
y.train <- Y[-Z]
y.test <- Y[Z]

library(class)
knn.result <- knn(x.train, x.test,y.train,3)
table(y.test, knn.result)

class.rate = rep(0,20)
for(k in 1:20){
  knn.result <- knn(x.train, x.test,y.train,k)
  class.rate[k] <- mean(y.test == knn.result)
}
which.max(class.rate)
plot(class.rate, type = "line", lwd = 4)
knn.in <- class.rate[which.max(class.rate)]
knn.in
```

Non parametric approach of KNN shows that the maximum prediction accuracy is obtained when k =19. It means that using 19 neighborhood to determine the vote of the classification, we are able to predict 79% of default and non default group correctly. 

```{r}
rf = randomForest(defaultnm ~ ., data=creditusd2, subset=-Z)
Yhat = predict(rf,creditusd2, type="class")
mean(Yhat[-Z] != creditusd2$defaultnm[-Z])

cv.err = rep(0,7)
n.trees= rep(0,7)
 for (m in 1:7){
rf.m = randomForest( defaultnm ~ ., data=creditusd2[-Z,], mtry=m )
opt.trees = which.min(rf.m$err.rate)
rf.m = randomForest( defaultnm ~ ., data=creditusd2[-Z,], mtry=m, ntree=opt.trees )
Yhat = predict(rf.m,newdata=creditusd2[Z,], type="class")
pred.err = mean( (Yhat == creditusd2$defaultnm[Z])^2 )
cv.err[m] = pred.err 
n.trees[m] = opt.trees
 }
which.min(cv.err)
n.trees[which.min(cv.err)]

rf.optimal = randomForest(defaultnm~., data = creditusd2, mtry = which.min(cv.err), ntree = n.trees[which.min(cv.err)])
rf.optimal
names(rf.optimal)
rf.optimal$confusion[2,3]
```



##conclusion

```{r conclusion,message=FALSE}

result <- tibble(InitialLogistic = log.in, 
VSLogistic = log.vs,
InitialLDA = lda.in,
VSLDA = lda.vs,
VSQDA = qda.vs,
InitialKNN = knn.in
#,randomForest = rf.optimal$confusion[2,3] 
)
result
```

##repeat the method using transformation data
#method1
```{r PC2,message=FALSE}
creditusd <- creditt[,c(1:12,25, 26, 28:39)]
cnumt2<- creditusd[,c(6:12,14:26)]
scnumt2 <- scale(cnumt2, scale = TRUE)
creditusd[,c(6:12,14:26)] <- scnumt2
creditusd <- creditusd %>% mutate(defaultnm = ifelse (defaultnm == "default", 1, 0))
creditusd$defaultnm <- as.numeric(creditusd$defaultnm)
creditusd
```


```{r machine learning2,message=FALSE}
#logistic Regression Initial model
creditusd <- creditusd[,-c(1:2)]
creditusd$defaultnm <- as.factor(creditusd$defaultnm)
mod1 <- glm(defaultnm~., data = creditusd, family = binomial(link="logit"))
summary(mod1)
#defaultnm~ limbalusd + SEX + MARRIAGE + AGE + PAY_0 + PAY_2 + PAY_3 +PAY_5 +PAY_6 + billamt6 + payamt1 + payamt2 + payamt3 + payamt4 + payamt5
creditusd

n <- length(creditusd$defaultnm)
Z <- sample(n,n/10)
c.test <- creditusd[Z,]
mod1.1 <- glm(defaultnm~ limbalusd + SEX + MARRIAGE + AGE + PAY_0 + PAY_2 + PAY_3 +PAY_5 +PAY_6 + billamt6 + payamt1 + payamt2 + payamt3 + payamt4 + payamt5 , data = creditusd[-Z,], family = binomial(link = "logit"))
summary(mod1.1)

prob = predict(mod1.1, data.frame(c.test))
yespred <- 1*(prob>0.5)

table(creditusd$defaultnm[Z], yespred)
log.in <- mean( creditusd$defaultnm[Z] == yespred)
log.in

attach(creditusd)

#TPR = rep(0,100)
#FPR = rep(0,100)
#for(k in 1:100){
#  fit = glm(defaultnm~ limbalusd + SEX + MARRIAGE + AGE + PAY_0 + PAY_2 + PAY_3 +PAY_5 +PAY_6 + billamt6 + payamt1 + payamt2 + payamt3 + payamt4 + payamt5 , data = creditusd[-Z,], family = binomial(link = "logit"))
#  prob  = predict(fit, data.frame(creditusd[Z,]), type = "response")
#  Yhat = 1*(prob > k/100)
#  TPR[k] = sum(Yhat ==1 & defaultnm==1)/ sum(defaultnm==1)
#  FPR[k] = sum(Yhat ==1 & defaultnm==0)/ sum(defaultnm==0)
#}
#plot(FPR, TPR, xlab="False positive rate", ylab="True positive rate", main="ROC curve")
#any kind of threshold will gave the same result



creditusd.prob.test <- predict(mod1.1, creditusd, type ="response")[-Z]
pred <- prediction(creditusd.prob.test,creditusd$defaultnm[-Z] )
perf <- performance(pred, "tpr","fpr")
plot(perf, colorize = TRUE)
auc=performance(pred, "auc") 
c(auc@y.name[[1]], auc@y.values[[1]])

```


```{r variable selection2,message=FALSE}


null = glm( defaultnm ~ 1, data=creditusd, family = binomial(link="logit") )
full = glm( defaultnm ~ ., data=creditusd, family = binomial(link="logit") )

#step( null, scope=list(lower=null, upper=full), direction="forward" )

#glm(formula = defaultnm ~ PAY_0 + payamt1 + PAY_4 + billamt6 + 
 #   payamt4 + MARRIAGE + EDUCATION + limbalusd + billamt2 + payamt3 + 
  #  PAY_6 + payamt2 + PAY_3 + SEX + billamt5 + payamt5 + PAY_5 + 
   # AGE + PAY_2, family = binomial(link = "logit"), data = creditusd)

#step( full, scope=list(lower=null, upper=full), direction="backward" )

#glm(formula = defaultnm ~ SEX + EDUCATION + MARRIAGE + AGE + 
 #   PAY_0 + PAY_2 + PAY_3 + PAY_4 + PAY_5 + PAY_6 + limbalusd + 
  #  billamt3 + billamt5 + billamt6 + payamt1 + payamt2 + payamt3 + 
   # payamt4 + payamt5, family = binomial(link = "logit"), data = creditusd)


modvs <- glm(formula = defaultnm ~ SEX + EDUCATION + MARRIAGE + AGE + 
PAY_0 + PAY_2 + PAY_3 + PAY_4 + PAY_5 + PAY_6 + limbalusd + 
  billamt3 + billamt5 + billamt6 + payamt1 + payamt2 + payamt3 + 
 payamt4 + payamt5, family = binomial(link = "logit"), data = creditusd)
summary(modvs)

#Education, pay_4, andpay_2 is not significant. 
#PAY_0 + payamt2 + payamt1 + PAY_3 + payamt6 + payamt5 + billamt3 + PAY_2 + billamt1

modvs2 <- glm(formula = defaultnm ~ SEX  + MARRIAGE + AGE + 
PAY_0  + PAY_3 + PAY_5 + PAY_6 + limbalusd + 
  billamt3 + billamt6 + payamt1 + payamt2 + payamt3 + 
 payamt4 + payamt5, family = binomial(link = "logit"), data = creditusd[-Z,])

prob = predict(modvs2, data.frame(c.test))
yespred <- 1*(prob>0.5)
table(creditusd$defaultnm[Z], yespred)
log.vs <- mean( creditusd$defaultnm[Z] == yespred)
log.vs
#logistic regression predict 78% correctly



#TPR = rep(0,100)
#FPR = rep(0,100)
#for(k in 1:100){
#  fit = glm(formula = defaultnm ~ SEX  + MARRIAGE + AGE + 
#PAY_0  + PAY_3 + PAY_5 + PAY_6 + limbalusd + 
#  billamt3 + billamt6 + payamt1 + payamt2 + payamt3 + 
# payamt4 + payamt5, family = binomial(link = "logit"), data = creditusd[-Z,])
 # prob  = predict(fit, data.frame(creditusd[Z,]), type = "response")
  #Yhat = 1*(prob > k/100)
  #TPR[k] = sum(Yhat ==1 & defaultnm==1)/ sum(defaultnm==1)
  #FPR[k] = sum(Yhat ==1 & defaultnm==0)/ sum(defaultnm==0)
#}
#plot(FPR, TPR, xlab="False positive rate", ylab="True positive rate", main="ROC curve")
#we can use any kind of threshold



creditusd.prob.test <- predict(modvs2, creditusd, type ="response")[-Z]
pred <- prediction(creditusd.prob.test,creditusd$defaultnm[-Z] )
perf <- performance(pred, "tpr","fpr")
plot(perf, colorize = TRUE)
auc=performance(pred, "auc") 
c(auc@y.name[[1]], auc@y.values[[1]])
```

##Linear Discriminant Analysis

```{r lda2,message=FALSE}

lda.fit <- lda(defaultnm~., data= creditusd, CV = TRUE)
table(creditusd$defaultnm, lda.fit$class)
lda.in <- mean(creditusd$defaultnm == lda.fit$class)
lda.in
```



```{r lda vs2,message=FALSE}
lda.fitvs <- lda( defaultnm ~ SEX + EDUCATION + MARRIAGE + AGE + 
 PAY_0 + PAY_2 + PAY_3 + PAY_4 + PAY_5 + PAY_6 + limbalusd + 
   billamt3 + billamt5 + billamt6 + payamt1 + payamt2 + payamt3 + 
  payamt4 + payamt5 , data = creditusd, CV = TRUE)
table(creditusd$defaultnm, lda.fitvs$class)
lda.vs <- mean(creditusd$defaultnm == lda.fitvs$class)
lda.vs
```

Both Linead Discriminant Analysis based on all variables and variable selection give us 81.1% of prediction accuracy. 

```{r qda2,message=FALSE}

###qda for all variable can't be done due to rank deficiency in one group. It means that the data has too many multicollinearity. 

qda.fitvs <- qda( defaultnm ~ limbalusd + SEX  + MARRIAGE + AGE + PAY_0 + PAY_2 + PAY_3 + PAY_5 + billamt1 + billamt2 +    billamt5 + payamt1 + payamt2 + payamt3 + payamt4 + payamt5 , data = creditusd, CV = TRUE)
table(defaultnm, qda.fitvs$class)
qda.vs<- mean(defaultnm == qda.fitvs$class,na.rm=TRUE)
qda.vs
```


```{r KNN2,message=FALSE}
creditusd2 <- creditusd
creditusd2[,3:5] <- credit2[,3:5]
creditusd2 <- creditusd2[,-1]
creditusd2$defaultnm <- as.factor(creditusd2$defaultnm)
n = length(creditusd2$defaultnm)
Z = sample( n, n/10)
X = model.matrix(defaultnm~., creditusd2)
Y = creditusd2$defaultnm
c.train = creditusd2[-Z,]
c.test = creditusd2[Z,]
x.train <- X[-Z,]
x.test <- X[Z,]
y.train <- Y[-Z]
y.test <- Y[Z]

library(class)
knn.result <- knn(x.train, x.test,y.train,3)
table(y.test, knn.result)

class.rate = rep(0,20)
for(k in 1:20){
  knn.result <- knn(x.train, x.test,y.train,k)
  class.rate[k] <- mean(y.test == knn.result)
}
which.max(class.rate)
plot(class.rate, type = "line", lwd = 4)
knn.in <- class.rate[which.max(class.rate)]
knn.in
```



```{r random forest2,message=FALSE}

rf = randomForest(defaultnm ~ ., data=creditusd2, subset=-Z)
Yhat = predict(rf,creditusd2, type="class")
mean(Yhat[-Z] != creditusd2$defaultnm[-Z])

cv.err = rep(0,7)
n.trees= rep(0,7)
 for (m in 1:7){
rf.m = randomForest( defaultnm ~ ., data=creditusd2[-Z,], mtry=m )
opt.trees = which.min(rf.m$err.rate)
rf.m = randomForest( defaultnm ~ ., data=creditusd2[-Z,], mtry=m, ntree=opt.trees )
Yhat = predict(rf.m,newdata=creditusd2[Z,], type="class")
pred.err = mean( (Yhat == creditusd2$defaultnm[Z])^2 )
cv.err[m] = pred.err 
n.trees[m] = opt.trees
 }
which.min(cv.err)
n.trees[which.min(cv.err)]

rf.optimal = randomForest(defaultnm~., data = creditusd2, mtry = which.min(cv.err), ntree = n.trees[which.min(cv.err)])
rf.optimal
names(rf.optimal)
rf.optimal$confusion[2,3]
```



##conclusion

```{r conclusion2,message=FALSE}

result2 <- tibble(InitialLogistic = log.in, 
VSLogistic2 = log.vs,
InitialLDA2 = lda.in,
VSLDA2 = lda.vs,
VSQDA2 = qda.vs,
InitialKNN2 = knn.in
,randomForest = rf.optimal$confusion[2,3] 
)
result
result2
```